Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings, Discriminative Training and Other Goodies

### 2010-04-23 papers/wmt10/joshua-progress.tex
done with variational decoding
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1615 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\documentclass[11pt]{article}
\usepackage{acl2010}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[colorlinks]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
%\setlength\titlebox{6.5cm}    % You can expand the title box if you
% really have to

\newcommand{\joshua}{\textbf{Joshua}}

@@

\section{Variational Decoding}
Statistical models in machine translation exhibit spurious ambiguity.
That is, the probability of an output string is split
among many distinct derivations (e.g., trees or
segmentations). In principle, the goodness of a
string is measured by the total probability of its
many derivations. However, finding the best string
(e.g., during decoding) is then computationally intractable.
Therefore, the first version implemented the Viterbi ?????Joshua most systems use a simple
Viterbi approximation that measures the goodness 
of a string using only its most probable derivation.

Instead, we develop a variational approximation \cite{variational-decoding-acl09},
which considers all the derivations but still
allows tractable decoding.

>>>>

\documentclass[11pt]{article}
\usepackage{acl2010}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[colorlinks]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xspace}

%\setlength\titlebox{6.5cm}    % You can expand the title box if you
% really have to

\newcommand{\joshua}{\textbf{Joshua}\xspace}
\newcommand{\Q}{\mathcal{Q}}

@@

\section{Variational Decoding}

Statistical models in machine translation exhibit spurious ambiguity.
That is, the probability of an output string is split
among many distinct derivations (e.g., trees or
segmentations). In principle, the goodness of a
string is measured by the total probability of its
many derivations. However, finding the best string
(e.g., during decoding) is then NP-hard.
The first version of \joshua implemented the
Viterbi approximation that measures the goodness 
of a string using only its most probable derivation.

The Viterbi approximation is efficient, but ignores most of the derivations in the hypergraph.
We implemented the variational decoding \cite{variational-decoding-acl09}, which works as follows.
Given a foreign string, the MT system first produces a hypergraph, 
which encodes a probability distribution $p$ over 
possible output strings and their derivations.
The variational decoding constructs a second distribution $q \in \Q$ 
that approximates $p$ as well as possible, 
and then finds the best string according to $q$ (instead of $p$).
The $q$ distribution is parameterized by an $n$-gram model, under which 
the last two steps can be performed efficiently via dynamic programming.
In this way, variational decoding considers all the derivations but still
allows tractable decoding.
